Deterministic Scheduling of Periodic Messages in a Network

### 2017-07-12 article/article.tex
article figures algo routes 0

 In this article, we study mechanisms to deal with the latency and periodicity constraints in the fronthaul network.  The current solution ~\cite{pizzinat2015things,tayq2017real} is to use use dedicated circuits, since statistical multiplexing even with a very large bandwidth, does not comply with the latency requirements.

 \todo{mettre la suite de l'état de l'art}

@@

 \todo{à revoir, citation ?}
 Thus in this article, we work on latency and periodicity constraints in fronthaul. The best current solution is to rely on an almost full optical approach, where each end-point (RRH on one side, BBU on the other side) is connected through direct fiber or full optical switches [7]. This architecture is very expensive and hardly scales in the case of a mobile network composed of about 10,000 base stations.

@@

      \todo{commencer les routes à 0 ou pas ?}

      A {\bf $(P,\tau)$-periodic assignment} of a routed network $(G,\cal R)$ is a sequence  ${\cal M}=(m_1, \ldots ,m_n)$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R$. The number $m_{i}$ represents the time at which a message is emitted at the first vertex of the route $r_{i}$. We will always consider that the $m_{i}$ are between $0$ and $P-1$. Moreover, \emph{no pair of routes must have a collision} in a $(P,\tau)$-periodic assignment.

@@

      From now on, we will only consider graphs with a very simple topology that we call the {\bf star topology}. 
      First, for each arc $(u,v)$, there is also an arc $(v,u)$ with the same weight.
      There are two sets of vertices, $S=\{s_1,...,s_n\}$ and $T=\{t_1,...,t_n\}$ of cardinality $n$ and two special nodes:
      the central source node {\bf $c_s$} and the central target node {\bf $c_t$}. There is an arc between {\bf $c_s$} and {\bf $c_t$} shared by all routes. For all $i$, there is an arc between $s_i$ and $c_s$ and between $t_i$ and $c_t$.
      The routes are the directed paths $r_i = s_i,c_s,c_t,t_i$ and $\rho(r_i) = t_i,c_t,c_s,s_i$ which defines a 
      symmetric routed network.

@@

\tikzset{
  LabelStyle/.style = { rectangle, rounded corners, draw,
                       font = \bfseries },
  EdgeStyle/.append style = {<->} }

  \SetGraphUnit{5}

  \node[draw,circle] (s3) at (0, 0) {$s_3$}; 
  \node[draw,circle] (s2) at (0, 2) {$s_2$}; 
  \node[draw,circle] (s1) at (0, 4) {$s_1$};

  \node[draw,circle] (t3) at (8, 0) {$t_3$}; 
  \node[draw,circle] (t2) at (8, 2) {$t_2$}; 
  \node[draw,circle] (t1) at (8, 4) {$t_1$};

@@

  \Edge[label = $\omega(s_1.c_s)$](s1)(cs)
  \Edge[label = $\omega(...)$](s2)(cs)
  \Edge[label = $\omega(...)$](s3)(cs)

@@

  When the waiting times are zero, $MT({\cal M})$ is equal to twice the weight of the longest route, thus $T_{max}$ is not relevant anymore. For a route $r$, choosing the offset $m_r$ also sets the offset of the route $\rho(r)$ $m_{\rho(r)} + \lambda(r) \mod P$.
  Consider an assignment $\{m_1,\dots,m_n\}$ of a star topology and let $m_i'= m_{i} - t(c_s,r_i) \mod P$.
  Then $\{m_1',\dots,m_n'\}$ is an assignment of the same star topology where the weights of the arcs from $s_i$ to $c_s$ are $0$
  for all $i$. Therefore we assume from now on that the \emph{weight of the arcs from $s_i$ to $c_s$ are all equal to $0$}.

@@

    We first present a simple policy, which works when the period is large with regard to the weight of the routes.
    The messages are sent in order from the shortest route to the longest route, without any gap between two messages in the forward period.
    In other words, we assume that the route $r_i$ are sorted by increasing $\lambda(r_i)$ and we set $m_{i}$ the offset of $r_i$ to $(i-1)\tau$. We call this algorithm {\bf Shortest-Longest}.

@@

      \begin{proposition} Let $(G, {\cal R})$ be a symmetric routed network, and let $n\tau + 2(\lambda(r_{n}) - \lambda(r_{1})) \leq P$, then there is $(P,\tau)$-periodic assignment of $(G, {\cal R})$ with all waiting times $0$ which is given by Shortest-Longest in time $O(n\log(n))$.\label{prop:SL}
      \end{proposition}
      \begin{proof}
       Since $m_{s_i} = (i-1)\tau$, $[t(c_s,r_{i})] = \{i\tau,\dots, (i+1)\tau -1\}$ and there are no collision on the forward period.

       We may assume that $\lambda(r_{1}) = 0$, since removing $\lambda(r_{1})$ from every arc $(c_t,t_i)$ does not change the order on the length of the routes nor the collisions between messages.
       Since $\lambda(r_{1}) = 0$, by hypothesis we have $n\tau + 2\lambda(r_{n}) \leq P$ which implies that
       $[t(c_s,r_{i})] = \{2 \lambda(r_{i}) + (i-1)\tau, \dots,  2 \lambda(r_{i}) + i\tau -1\}$.
       Since $ \lambda(r_{i}) \leq  \lambda(r_{i+1})$ by construction, we have  $2 \lambda(r_{i}) + i\tau -1 < 2 \lambda(r_{i+1}) + i\tau$ which proves that there are no collision on the backward period.

@@

     \begin{algorithm}[H]
    \caption{ Greedy deadline ({\bf GD}) }
    \begin{algorithmic}
    \REQUIRE A routed network $(G,{\cal R})$,a period $P$, packet size $\tau$, $ T_{max}$, the offsets $m_i$
    \ENSURE $(P-\tau)-$periodic assignment of $(G,{\cal R})$, of failure
   \STATE  ${\cal H} \leftarrow$ empty binary min heap
       \STATE  free\_ intervals $\leftarrow$ [0,$P$]

@@

      \WHILE{There is some non-assigned routes}
      \IF{${\cal H}$ is empty}
      \STATE t $\leftarrow $ min\_non\_assigned(eligible\_time)
      \STATE take $r_i$ the route with eligible\_time[$r_i$] = t
      \STATE ${\cal H}$ $\leftarrow$ insert ($r_i$, deadlines($r_i$)).
      \ENDIF

      \STATE Extract the route $r_i$ with the lowest deadline from ${\cal H}$
      \STATE t $\leftarrow$ beginning of the first free interval of size $\tau$ after t
      \STATE $w_i \leftarrow$ t - eligible\_time[$r_i$]
      \STATE update(free\_ intervals)
      \STATE t $\leftarrow$ t + $\tau$
      \FORALL{routes $r_i$ with  eligible\_time[$r_i$] $\leq$ t}
      \STATE ${\cal H}$ $\leftarrow$ insert ($r_i$, deadlines($r_i$).
      \ENDFOR
      \ENDWHILE
%    \STATE  $w_i \leftarrow 0$
%    \FORALL{route $r_{t_i}$}
%    \STATE  $w_i \leftarrow 0$
%    \STATE period-end $\leftarrow m_{s_i} + \lambda(r_{s_i}) + t(c_t,r_{t_i}) + P$
%    \FORALL{route $r_{t_j}$}
%    \STATE deadline-route$ \leftarrow m_{s_j} + T_{max}-t(c_s,r_{s_j})$
%    \STATE $deadline \leftarrow$ min(deadline-route,period-end)
%    \ENDFOR

    \end{algorithmic}
    \end{algorithm}
    The function  min\_non\_assigned(eligible\_time) returns the lowest time in eligible\_time, for which the corresponding route is not assigned yet. The function update(free\_ intervals) removes an interval of size $\tau$, which correspond to the message,  from free\_ intervals.

>>>>

 In this article, we study mechanisms to deal with the latency and periodicity constraints in the fronthaul network.  The current solution ~\cite{pizzinat2015things} is to use use dedicated circuits, since statistical multiplexing even with a very large bandwidth, does not comply with the latency requirements. 
 \todo{mettre la suite de l'Etat de l'art}
 Thus in this article, we work on latency and periodicity constraints in fronthaul. The best current solution is to rely on an almost full optical approach, where each end-point (RRH on one side, BBU on the other side) is connected through direct fiber or full optical switches ~\cite{tayq2017real}. This architecture is very expensive and hardly scales in the case of a mobile network composed of about 10,000 base stations.

@@

@@

      \todo{commencer les routes  0 ou pas ?}

      A {\bf $(P,\tau)$-periodic assignment} of a routed network $(G,\cal R)$ is a sequence  ${\cal M}=(m_0, \ldots ,m_{n-1})$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R$. The number $m_{i}$ represents the time at which a message is emitted at the first vertex of the route $r_{i}$. We will always consider that the $m_{i}$ are between $0$ and $P-1$. Moreover, \emph{no pair of routes must have a collision} in a $(P,\tau)$-periodic assignment.

@@

      From now on, we will only consider graphs with a very simple topology that we call the {\bf star topology}. 
      First, for each arc $(u,v)$, there is also an arc $(v,u)$ with the same weight.
      There are two sets of vertices, $S=\{s_0,...,s_{n-1}\}$ and $T=\{t_0,...,t_{n-1}\}$ of cardinality $n$ and two special nodes:
      the central source node {\bf $c_s$} and the central target node {\bf $c_t$}. There is an arc between {\bf $c_s$} and {\bf $c_t$} shared by all routes. For all $i$, there is an arc between $s_i$ and $c_s$ and between $t_i$ and $c_t$.
      The routes are the directed paths $r_i = s_i,c_s,c_t,t_i$ and $\rho(r_i) = t_i,c_t,c_s,s_i$ which defines a 
      symmetric routed network.

@@

\tikzset{EdgeStyle/.style={<->,font=\scriptsize,above,sloped,midway}}
  \SetGraphUnit{5}

  \node[draw,circle] (s3) at (0, 0) {$s_2$}; 
  \node[draw,circle] (s2) at (0, 2) {$s_1$}; 
  \node[draw,circle] (s1) at (0, 4) {$s_0$};

  \node[draw,circle] (t3) at (8, 0) {$t_2$}; 
  \node[draw,circle] (t2) at (8, 2) {$t_1$}; 
  \node[draw,circle] (t1) at (8, 4) {$t_0$};

@@

  \Edge[label = $\omega(s_1\,c_s)$](s1)(cs)
  \Edge[label = $\omega(...)$](s2)(cs)
  \Edge[label = $\omega(...)$](s3)(cs)

@@

  When the waiting times are zero, $MT({\cal M})$ is equal to twice the weight of the longest route, thus $T_{max}$ is not relevant anymore. For a route $r$, choosing the offset $m_r$ also sets the offset of the route $\rho(r)$ $m_{\rho(r)} + \lambda(r) \mod P$.
  Consider an assignment $\{m_0,\dots,m_{n-1}\}$ of a star topology and let $m_i'= m_{i} - t(c_s,r_i) \mod P$.
  Then $\{m_0',\dots,m_{n-1}'\}$ is an assignment of the same star topology where the weights of the arcs from $s_i$ to $c_s$ are $0$
  for all $i$. Therefore we assume from now on that the \emph{weight of the arcs from $s_i$ to $c_s$ are all equal to $0$}.

@@

    We first present a simple policy, which works when the period is large with regard to the weight of the routes.
    The messages are sent in order from the shortest route to the longest route, without any gap between two messages in the forward period.
    In other words, we assume that the route $r_i$ are sorted by increasing $\lambda(r_i)$ and we set $m_{i}$ the offset of $r_i$ to $i\tau$. We call this algorithm {\bf Shortest-Longest}.

@@

      \begin{proposition} Let $(G, {\cal R})$ be a symmetric routed network, and let $n\tau + 2(\lambda(r_{n}) - \lambda(r_{0})) \leq P$, then there is $(P,\tau)$-periodic assignment of $(G, {\cal R})$ with all waiting times $0$ which is given by Shortest-Longest in time $O(n\log(n))$.\label{prop:SL}
      \end{proposition}
      \begin{proof}
       Since $m_{s_i} = i\tau$, $[t(c_s,r_{i})] = \{i\tau,\dots, (i+1)\tau -1\}$ and there are no collision on the forward period.

       We may assume that $\lambda(r_{0}) = 0$, since removing $\lambda(r_{0})$ from every arc $(c_t,t_i)$ does not change the order on the length of the routes nor the collisions between messages.
       Since $\lambda(r_{0}) = 0$, by hypothesis we have $n\tau + 2\lambda(r_{n}) \leq P$ which implies that
       $[t(c_t,r_{i})] = \{2 \lambda(r_{i}) + i\tau, \dots,  2 \lambda(r_{i}) + (i+1)\tau -1\}$.
       Since $ \lambda(r_{i}) \leq  \lambda(r_{i+1})$ by construction, we have  $2 \lambda(r_{i}) + i\tau -1 < 2 \lambda(r_{i+1}) + (i+1)\tau$ which proves that there are no collision on the backward period.

@@

     \begin{algorithm}[H]
    \caption{ Greedy deadline ({\bf GD}) }
    \begin{algorithmic}
    \REQUIRE A routed network $(G,{\cal R})$, a period $P$, packet size $\tau$, $ T_{max}$, the offsets $m_i$
    \ENSURE $(P,\tau)$-periodic assignment of $(G,{\cal R})$, or failure
   \STATE  ${\cal H} \leftarrow$ empty set //{\em set of eligible routes}
       \STATE  free\_ intervals $\leftarrow$ [0,$P$] //{\em list of intervals of free slots}

@@

      \WHILE{There is some non-assigned routes}
      \IF{${\cal H}$ is empty}
      \STATE t $\leftarrow $ min\_non\_assigned(eligible\_time)
      \STATE take $r_i$ the route with eligible\_time[$r_i$] = t
      \STATE 
      \ENDIF

      \STATE $r \leftarrow $ extract\_min(${\cal H}$)
      \STATE t $\leftarrow$ next\_free\_interval(free\_intervals, t) //{\em if there is no more free intervals of size $\tau$, the algorithm fails}
      \STATE $w_i \leftarrow$ t - eligible\_time[$r_i$]
      \STATE update(t,free\_ intervals)
      \STATE t $\leftarrow$ t + $\tau$
      \FORALL{routes $r_i$ with  eligible\_time[$r_i$] $\leq$ t}
      \STATE insert ( ${\cal H}$,$r_i$).
      \ENDFOR
      \ENDWHILE
%    \STATE  $w_i \leftarrow 0$
%    \FORALL{route $r_{t_i}$}
%    \STATE  $w_i \leftarrow 0$
%    \STATE period-end $\leftarrow m_{s_i} + \lambda(r_{s_i}) + t(c_t,r_{t_i}) + P$
%    \FORALL{route $r_{t_j}$}
%    \STATE deadline-route$ \leftarrow m_{s_j} + T_{max}-t(c_s,r_{s_j})$
%    \STATE $deadline \leftarrow$ min(deadline-route,period-end)
%    \ENDFOR

    \end{algorithmic}
    \end{algorithm}
    The function  min\_non\_assigned(eligible\_time) returns the lowest time in eligible\_time, for which the corresponding route is not assigned yet. The function update(free\_intervals) removes an interval of size $\tau$ beginning at t, which correspond to the message,  from free\_intervals.
